---
title: "Permutation test vignette: WPP 2024 development groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Permutation test vignette: WPP 2024 development groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6) 
```

```{r setup}
library(dplyr)
library(readr)
library(testthat)
```

This vignette serves as an example of how to use the implemented functions to perform a two-group permutation test for comparing demographic indicators across development groups using country-level data.

Specifically, the following examples demonstrate:

-   two-group permutation test based on the difference in means
    `perm_test_two_group()`,

-   computation of permutation-based p-values
    `perm_p_value()`, and

-   visualization of permutation distributions with the observed statistic
    `plot_perm_dist()`.

In particular, the examples focus on country-level estimates from the World Population Prospects 2024 dataset and illustrate how permutation-based inference can be used to assess whether observed differences between development groups are statistically significant.

# Loading the Functions

```{r}
source("data_preparation.R")
source("permutation_functions.R")
source("plot_perm_dist.R")
```


# Loading the Cleaned Dataset

We use the cleaned country-level dataset derived from the World Population Prospects 2024 Estimates data.

```{r}
wpp_path <- "../Data/wpp_country_indicators_2023.csv"
wpp <- readr::read_csv(wpp_path, show_col_types = FALSE)

dplyr::glimpse(wpp)
```

# Preparing Data for a Selected Indicator

In this vignette we demonstrate the workflow using one indicator (birth rate). The same code applies to life expectancy or mortality rate by changing `indicator_col`.

```{r}
test_df <- perm_data_extract(wpp, birth_rate, dev_group)

dplyr::count(test_df, group)
summary(test_df$outcome)
```

# Running the Permutation Test

We run a two-sided permutation test using the difference in means (More developed - Less developed) as the test statistic.

## Hypotheses

Let $\mu_M$ be the mean outcome for **More developed** and $\mu_L$ for **Less developed**.

$$
H_0: \mu_M = \mu_L
$$

$$
H_A: \mu_M \ne \mu_L
$$

We use the observed test statistic:

$$
T_{\text{obs}} = \bar{X}_M - \bar{X}_L
$$

and generate a permutation distribution by shuffling the group labels $B$ times.

```{r}
res <- perm_test_two_group_mean(
  df = test_df,
  B = 2000,
  seed = 1,
  alternative = "two.sided"
)

res$observed
res$p_value
```

# Visualizing the Permutation Distribution

We plot the permutation distribution and highlight the observed test statistic with a vertical line.

```{r}
plot_perm_dist(res, bins = 40)
```


# Inspecting Outcome Distributions by Group

```{r}
par(mfrow = c(1, 2))

hist(
  test_df$outcome[test_df$group == "More developed"],
  main = "More developed: outcome",
  xlab = "Outcome",
  breaks = 30
)

hist(
  test_df$outcome[test_df$group == "Less developed"],
  main = "Less developed: outcome",
  xlab = "Outcome",
  breaks = 30
)

par(mfrow = c(1, 1))
```

# Conclusion (reject vs. fail to reject)

We conclude using a significance level α=0.05.

```{r results="asis"}
alpha <- 0.05

if (res$p_value < alpha) {
  cat("Since p <", alpha, ", we reject $H_0$.\n\n")
  cat("The observed difference in means is unlikely under random label assignment, ",
      "providing statistical evidence that the indicator differs between development groups in 2023.\n")
} else {
  cat("Since p \\ge", alpha, ", we fail to reject $H_0$.\n\n")
  cat("The observed difference in means is consistent with the permutation distribution ",
      "under exchangeability; there is insufficient statistical evidence of a difference between development groups.\n")
}
```

# Bibliography

