---
title: "Permutation test vignette: WPP 2024 development groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Permutation test vignette: WPP 2024 development groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=7, fig.height=6) 
```

```{r setup}
library(dplyr)
library(readr)
library(testthat)
```

This vignette demonstrates how to use the package functions to perform a two-group Monte Carlo permutation test comparing demographic indicators across development groups using country-level data.

The workflow shown includes:

- preparing a clean two-column dataset with `perm_data_extract()`,

- computing a two-sample Kolmogorov–Smirnov (KS) statistic with `perm_stat_ks()`,

- running a permutation test with `perm_test_two_group_ks()`, and

- visualizing the permutation distribution with `plot_perm_dist()`.

# Loading the Functions

```{r}
source("data_preparation.R")
source("permutation_functions.R")
source("plot_perm_dist.R")
```

# Loading the Cleaned Dataset

We use the cleaned country-level dataset derived from the World Population Prospects 2024 Estimates data.

```{r}
wpp_path <- "../Data/wpp_country_indicators_2023.csv"
wpp <- readr::read_csv(wpp_path, show_col_types = FALSE)

dplyr::glimpse(wpp)
```

# Preparing Data for a Selected Indicator

In this vignette we demonstrate the workflow using one indicator (birth rate). The same code applies to life expectancy or mortality rate by changing `indicator_col`.

```{r}
test_df <- perm_data_extract(wpp, birth_rate, dev_group)

dplyr::count(test_df, group)
summary(test_df$outcome)
```

# Running the Permutation Test (KS statistic)

We run a two-sided permutation test using the KS statistic, which measures the maximum difference between the empirical CDFs of the two groups.

## Hypotheses

Let $F_M(x)$ be the distribution of the **More developed** group and $F_L(x)$ be the distribution of the **Less developed** group.

$$H_0: F_M(x) = F_L(x)$$
$$H_A: F_M(x) \neq F_L(x)$$

The observed test statistic is the Kolmogorov-Smirnov distance, defined as:

$$D_{obs} = \sup_x |F_M(x) - F_L(x)|$$
We generate a permutation distribution by shuffling the group labels 𝐵 times.

```{r}
res <- perm_test_two_group_ks(
  df = test_df,
  B = 2000,
  seed = 1,
  alternative = "two.sided"
)

res$observed
res$p_value
```

## Notes on the p-value

The function computes a Monte Carlo permutation p-value using the standard adjustment:

$p = \frac{\#\{\text{extreme permutations}\} + 1}{B + 1}$

This prevents reporting a p-value of exactly 0 when 𝐵is finite.


# Visualizing the Permutation Distribution

We plot the permutation distribution and highlight the observed KS statistic with a vertical line.

```{r}
plot_perm_dist(res, bins = 40)
```

Interpretation: if the observed statistic lies far in the tail of the permutation distribution, then the observed group difference is unlikely under label exchangeability.

# Inspecting Outcome Distributions by Group

```{r}
par(mfrow = c(1, 2))

hist(
  test_df$outcome[test_df$group == "More developed"],
  main = "More developed: outcome",
  xlab = "Outcome",
  breaks = 30
)

hist(
  test_df$outcome[test_df$group == "Less developed"],
  main = "Less developed: outcome",
  xlab = "Outcome",
  breaks = 30
)

par(mfrow = c(1, 1))
```

# Conclusion (reject vs. fail to reject)

We conclude using a significance level α=0.05.

```{r results="asis"}
alpha <- 0.05

if (res$p_value < alpha) {
  cat("Since p <", alpha, ", we reject $H_0$.\n")
  cat("The observed KS statistic is unlikely under random label assignment, ")
  cat("providing statistical evidence that the indicator distribution differs between development groups in 2023.\n")
} else {
  cat("Since p \\ge", alpha, ", we fail to reject $H_0$.\n")
  cat("The observed KS statistic is consistent with the permutation distribution under exchangeability; ")
  cat("there is insufficient statistical evidence of a distributional difference between development groups.\n")
}
```

# Generative AI Usage Statement

-   **`permutation_functions.R`** Clarifying roxygen documentation formatting and understanding tidy evaluation syntax (using {{ }}) within the wrapper functions.

-   **`plot_perm_dist.R`** Learning the usage of `invisible()` and `rgb()` functions.

-   **`test-perm_function.R`** ChatGPT was used to suggest reasonable unit test cases and to explain the use of `skip_if_not()` for conditionally skipping wrapper tests when `perm_data_extract` is not yet implemented. Some test functions are named by the suggestion of AI.

# Bibliography
