---
title: "Permutation test vignette: WPP 2024 development groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Permutation test vignette: WPP 2024 development groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib
nocite: "@*"
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 6
)

library(Sta380Project)
library(dplyr)
library(readr)
```

This vignette demonstrates how to use the package functions to perform a two-group Monte Carlo permutation test comparing demographic indicators across development groups using country-level data.

The workflow shown includes:

- preparing a clean two-column dataset with `perm_data_extract()`,

- computing a two-sample Kolmogorov–Smirnov (KS) statistic with `perm_stat_ks()`,

- running a permutation test with `perm_test_two_group_ks()`, and

- visualizing the permutation distribution with `plot_perm_dist()`.

# Loading the Cleaned Dataset

We use the cleaned country-level dataset derived from the World Population Prospects 2024 Estimates data.

```{r}
# wpp_path <- "../Data/wpp_country_indicators_2023.csv"
# wpp <- readr::read_csv(wpp_path, show_col_types = FALSE)
data("wpp", package = "Sta380Project")

dplyr::glimpse(wpp)
```

# Preparing Data for a Selected Indicator

In this vignette we demonstrate the workflow using one indicator (birth rate). The same code applies to life expectancy or mortality rate by changing `indicator_col`.

```{r}
test_df <- perm_data_extract(wpp, birth_rate, dev_group)

dplyr::count(test_df, group)
summary(test_df$outcome)
```

# Running the Permutation Test (KS statistic)

We run a two-sided permutation test using the KS statistic, which measures the maximum difference between the empirical CDFs of the two groups.

## Hypotheses

Let $F_M(x)$ be the distribution of the **More developed** group and $F_L(x)$ be the distribution of the **Less developed** group.

$$H_0: F_M(x) = F_L(x)$$
$$H_A: F_M(x) \neq F_L(x)$$

The observed test statistic is the Kolmogorov-Smirnov distance, defined as:

$$D_{obs} = \sup_x |F_M(x) - F_L(x)|$$
We generate a permutation distribution by shuffling the group labels 𝐵 times.

```{r}
res <- perm_test_two_group_ks(
  df = test_df,
  B = 2000,
  seed = 1,
  alternative = "two.sided"
)

res$observed
res$p_value
```

## Notes on the p-value

The function computes a Monte Carlo permutation p-value using the standard adjustment:

$p = \frac{\#\{\text{extreme permutations}\} + 1}{B + 1}$

This prevents reporting a p-value of exactly 0 when 𝐵is finite.


# Visualizing the Permutation Distribution

We plot the permutation distribution and highlight the observed KS statistic with a vertical line.

```{r}
plot_perm_dist(res, bins = 40)
```

Interpretation: The observed KS statistic lies far in the right tail of the permutation distribution, indicating that the observed difference between the two groups is much larger than would typically arise under random label assignment.

# Inspecting Outcome Distributions by Group

```{r}

outcome <- test_df$outcome
group <- test_df$group

groups <- as.character(group)
groups <- unique(groups)
  
if (length(groups) != 2) {
    stop("group must have exactly two levels. ")
}
xA <- outcome[group == groups[1]]
xB <- outcome[group == groups[2]]
  
ecdf_A <- stats::ecdf(xA)
ecdf_B <- stats::ecdf(xB)
  
plot(ecdf_A, col="blue", lwd=1,
       main="ECDF Comparison between two groups",
       xlab="Outcome",
       ylab="Cumulative Probability")
  
lines(ecdf_B, col="yellow", lwd=1)
  
legend("bottomright",
         legend=c("More developed","Less developed"),
         col=c("blue","yellow"),
         lwd=2)
```

# Conclusion (reject vs. fail to reject)

We conclude using a significance level α=0.05.

```{r results="asis"}
alpha <- 0.05
p_value <- round(res$p_value, 4)

perm_test_conclusion(p_value, alpha)
```

# Generative AI Usage Statement

-   **`permutation_functions.R`** Clarifying roxygen documentation formatting and understanding tidy evaluation syntax (using {{ }}) within the wrapper functions.

-   **`plot_perm_dist.R`** Learning the usage of `invisible()` and `rgb()` functions.

-   **`test-perm_function.R`** ChatGPT was used to suggest reasonable unit test cases and to explain the use of `skip_if_not()` for conditionally skipping wrapper tests when `perm_data_extract` is not yet implemented. Some test functions are named by the suggestion of AI.

# Bibliography